---
title: "CellChat"
author: "Caden McQuillen"
date: '2023-01-30'
output: 
  html_document: 
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Extract data from Seurat object and add new cell type vector
```{r Seurat Extraction, eval=FALSE}
library(Seurat)
PRN_6wk_8wk = readRDS("./Seurat_Objects/Annotated_PRN_6wk_8wk.rds")

#add 2nd cell type vector with separate Pou2f3 Neuroendocrine cluster
PRN_6wk_8wk$cell.type2 <- PRN_6wk_8wk$seurat_clusters
PRN_6wk_8wk$cell.type2 <- plyr::mapvalues(
    x = PRN_6wk_8wk$seurat_clusters,
    from = c("0","1","2","3","4","5","6","7","8","9","10","11","12","13","14","15","16"),
    to = c("Granulocytes","Macrophages","Luminal Epithelium","Fibroblasts","Basal Epithelium","Fibroblasts","Basal Epithelium","Neuroendocrine","Fibroblasts","Macrophages","Seminal Vesicle","Macrophages",    "Macrophages","Endothelium","Pou2f3_Neuroendocrine","Luminal Epithelium",   "Fibroblasts"))
```

## Extract data from Seurat object
```{r Suerat Extraction, eval=FALSE}
library(Seurat)
PRN_6wk_8wk = readRDS("./Seurat_Objects/Annotated_PRN_6wk_8wk.rds")
data.input <- GetAssayData(PRN_6wk_8wk, assay = "RNA", slot = "data") # normalized data matrix
labels <- PRN_6wk_8wk$cell.type3
meta <- data.frame(group = labels, row.names = names(labels)) # create a dataframe of the cell labels
```

## Running CellChat Part 1
```{r CellChat P1, eval=TRUE}
library(CellChat)
library(patchwork)
options(stringsAsFactors = FALSE)

#create cellchat object
cellchat <- createCellChat(object = data.input, meta = meta, group.by = "group")

#select DB (we are using mouse)
CellChatDB <- CellChatDB.mouse

#Use entire DB
CellChatDB.use <- CellChatDB # simply use the default CellChatDB

# set the used database in the object
cellchat@DB <- CellChatDB.use

# subset the expression data of signaling genes for saving computation cost
cellchat <- subsetData(cellchat) # This step is necessary even if using the whole database
future::plan("multiprocess", workers = 4) # do parallel
cellchat <- identifyOverExpressedGenes(cellchat)
cellchat <- identifyOverExpressedInteractions(cellchat)

# project gene expression data onto PPI (Optional)
cellchat <- projectData(cellchat, PPI.mouse)

saveRDS(cellchat, "PRN_6wk_8wk_cellchat_part1.rds")
```

## Running CellChat Part 2
```{r CellChat Part 2, eval=TRUE}

#triMean is used for calculating the average gene expression per cell group. 
cellchat <- computeCommunProb(cellchat)

# Filter out the cell-cell communication if there are only few number of cells in certain cell groups
cellchat <- filterCommunication(cellchat, min.cells = 10)


# data frame consisting of all the inferred cell-cell communications at the level of ligands/receptors
df.net_ligand_receptor <- subsetCommunication(cellchat)
write.csv(df.net_ligand_receptor, "Ligand_Receptor_CellChat_df.csv")

#compute the communication probability on signaling pathway level by summarizing the communication probabilities of all ligands-receptors interactions associated with each signaling pathway
cellchat <- computeCommunProbPathway(cellchat)
df.net_pathway <- subsetCommunication(cellchat, slot.name = "netP")
write.csv(df.net_pathway, "Pathway_CellChat_df.csv")

#calculate the aggregated cell-cell communication network by counting the number of links or summarizing the communication probability.
cellchat <- aggregateNet(cellchat)

#visualize the aggregated cell-cell communication network
groupSize <- as.numeric(table(cellchat@idents))
par(mfrow = c(1,2), xpd=TRUE)
netVisual_circle(cellchat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions")
netVisual_circle(cellchat@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Interaction weights/strength")

saveRDS(cellchat, "PRN_6wk_8wk_CellChat_Inferred_network.rds")
```
```{r Cell by Cell visualization, eval=FALSE}
#visualize the aggregated cell-cell communication network cell by cell
mat <- cellchat@net$weight
par(mfrow = c(3,4), xpd=TRUE)
for (i in 1:nrow(mat)) {
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  mat2[i, ] <- mat[i, ]
  plot = netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat), title.name = rownames(mat)[i])
}
 print(plot)
 
 
```

## Network Interaction Comparisons Pou2f3+ vs Pou2f3-

```{r, eval=TRUE}
#Extract Signaling pathways where Pou2f3+ and Pou2f3- are targets
Pou2f3_Pos_Target = subset(df.net_pathway, df.net_pathway$target == "Pou2f3_Neuroendocrine")
Pou2f3_Neg_Target = subset(df.net_pathway, df.net_pathway$target == "Neuroendocrine")

#Extract Signaling pathways where Pou2f3+ and Pou2f3- are sources
Pou2f3_Pos_Source = subset(df.net_pathway, df.net_pathway$source == "Pou2f3_Neuroendocrine")
Pou2f3_Neg_Source = subset(df.net_pathway, df.net_pathway$source == "Neuroendocrine")

#Pou2f3- Target: Unique Signaling Pathways
#Source: Basal Epithelial: SEMA3, WNT
#        Endothelium: SEMA6        
#        Fibroblasts: EPHB, IGF, NCAM, WNT, ncWNT
#        Luminal Epithelium: EPHB, SEMA3, WNT




#Pou2f3- to Pou2f3- : CADM, CDH, EPHA, GRN, JAM, KIT, L1CAM, MPZ, NCAM, OCLN, PTN, THBS
#Pou2f3+ to Pou2f3- : BMP, CADM, CDH, EPHB, GRN, JAM, MPZ, OCLN




#Pou2f3+ Target: Unique Signaling Pathways
#Source: Luminal Epithelium: EPHA

#Pou2f3- Source: Unique Signaling Pathways
#Target: Basal Epithelium: EPHA, PTN
#        Endothelium: CALCR, EPHA, NCAM, PTN, THBS
#        Fibroblasts: EPHA, NCAM, NRXN, NT, PDGF, PTN,THBS
#        Grunolcytes: PTN, THBS
#        Luminal Epithelium: EPHA
#        Macrophages:	PTN, THBS
#        Seminal Vesicle: PTN



#Pou2f3+ Source: Unique Signaling Pathways
#Target: Basal Epithelium: BMP, EPHB, MIF
#        Endothelium: EPHB
#        Fibroblasts: BMP, EPHB
#        Luminal Epithelium: BMP
#        Macrophages: IL10, MIF









```
## Pou2f3- Target: Unique Signaling Pathways
Source:\
Basal Epithelium: **SEMA3**, **WNT**\
Endothelium: **SEMA6**\
Fibroblasts: **EPHB**, **IGF**, **NCAM**, **WNT**, **ncWNT**\
Luminal Epithelium: **EPHB**, **SEMA3**, **WNT**\

## Pou2f3+ Target: Unique Signaling Pathways
Source:\
Luminal Epithelium: **EPHA**\
        
        
## Pou2f3- Source: Unique Signaling Pathways
Target:\
Basal Epithelium: **EPHA**, **PTN**\
Endothelium: **CALCR**, **EPHA**, **NCAM**, **PTN**, **THBS**\
Fibroblasts: **EPHA**, **NCAM**, **NRXN**, **NT**, **PDGF**, **PTN**,**THBS**\
Granulocytes: **PTN**, **THBS**\
Luminal Epithelium: **EPHA**\
Macrophages:	**PTN**, **THBS**\
Seminal Vesicle: **PTN**\



## Pou2f3+ Source: Unique Signaling Pathways
Target:\
Basal Epithelium: **BMP**, **EPHB**, **MIF**\
Endothelium: **EPHB**\
Fibroblasts: **BMP**, **EPHB**\
Luminal Epithelium: **BMP**\
Macrophages: **IL10**, **MIF**\

## Pou2f3- to Pou2f3+  
CADM, CDH, EPHA, GRN, JAM, KIT, L1CAM, MPZ, NCAM, OCLN, PTN, THBS

## Pou2f3+ to Pou2f3-  
BMP, CADM, CDH, EPHB, GRN, JAM, MPZ, OCLN

## Heatmap of EPHB and EPHA
```{r, eval = FALSE}
# Heatmap 
par(mfrow=c(1,1))
netVisual_heatmap(cellchat, signaling = "EPHA", color.heatmap = "Reds")
netVisual_heatmap(cellchat, signaling = "EPHB", color.heatmap = "Reds")
netVisual_heatmap(cellchat, signaling = "IL10", color.heatmap = "Reds")
netVisual_heatmap(cellchat, signaling = "MIF", color.heatmap = "Reds")
netVisual_heatmap(cellchat, signaling = "BMP", color.heatmap = "Reds")
netVisual_heatmap(cellchat, signaling = "THBS", color.heatmap = "Reds")
netVisual_heatmap(cellchat, signaling = "WNT", color.heatmap = "Reds")
netVisual_heatmap(cellchat, signaling = "PTN", color.heatmap = "Reds")
netVisual_heatmap(cellchat, signaling = "KIT", color.heatmap = "Reds")
#Ligand Receptor pair contribution to IL10
netAnalysis_contribution(cellchat, signaling = "IL10")
#Gene expression
plotGeneExpression(cellchat, signaling = "IL10")


#Ligand Receptor pair contribution to IL10
netAnalysis_contribution(cellchat, signaling = "WNT")
#Gene expression
plotGeneExpression(cellchat, signaling = "WNT")

#Ligand Receptor pair contribution to IL10
netAnalysis_contribution(cellchat, signaling = "KIT")
#Gene expression
plotGeneExpression(cellchat, signaling = "KIT")


```

## Signaling roles

```{r}
cellchat <- netAnalysis_computeCentrality(cellchat, slot.name = "netP") # the slot 'netP' means the inferred intercellular communication network of signaling pathways
# Visualize the computed centrality scores using heatmap, allowing ready identification of major signaling roles of cell groups
netAnalysis_signalingRole_network(cellchat, signaling = "EPHA", width = 8, height = 2.5, font.size = 10)
netAnalysis_signalingRole_network(cellchat, signaling = "EPHB", width = 8, height = 2.5, font.size = 10)
```

## Global Patterns
```{r}
library(NMF)
library(ggalluvial)
selectK(cellchat, pattern = "outgoing")
```
```{r}
nPatterns = 3
cellchat <- identifyCommunicationPatterns(cellchat, pattern = "outgoing", k = nPatterns)
netAnalysis_river(cellchat, pattern = "outgoing")
```

