---
title: "Sawyers_data"
author: "Caden McQuillen"
date: '2023-05-08'
output: html_document
---

## Convert Anndata object to Seurat
```{r}
library(Seurat)
library(SeuratData)
library(SeuratDisk)
library(MuDataSeurat)


sawyers <- ReadH5AD("./Sawyers_data/GSM6428979_adata.gemm.wt_plus_gfp.final.h5ad", layers = c("counts", "data"))


```


```{r}
DefaultAssay(sawyers) <- "RNA"
sawyers <- SetIdent(sawyers, value = "cell_type_final")
DimPlot(sawyers, reduction = "umap")
FeaturePlot(object = sawyers, features = 'POU2F3', slot = "scale.data")
```

## Pseudobullk of Sawyers Pou2f3 cluster
```{r}

sawyers_pou2f3 <- subset(sawyers, idents = "POU2F3")
sawyers_pou2f3 <- rowMeans(sawyers_pou2f3@assays$RNA@data)

names(sawyers_pou2f3)  <- tolower(names(sawyers_pou2f3))

```

## Pseudobulk of PRN Pou2f3
```{r}
PRN <- readRDS("./Seurat_Objects/Annotated_PRN_6wk_8wk.rds")
PRN <- SetIdent(PRN, value = "cell.type3")
PRN_pou2f3 <- subset(PRN, idents = "Pou2f3_Neuroendocrine")
rm(PRN)

PRN_pou2f3 <- rowMeans(PRN_pou2f3@assays$RNA@counts)
names(PRN_pou2f3)  <- tolower(names(PRN_pou2f3))

```

## Spearmans correlation of Pou2f3 clusters
```{r}
overlap <- intersect(names(sawyers_pou2f3), names(PRN_pou2f3))
sawyers_pou2f3 <- sawyers_pou2f3[names(sawyers_pou2f3) %in% overlap]
PRN_pou2f3 <- PRN_pou2f3[names(PRN_pou2f3) %in% overlap]

sawyers_pou2f3 <- sawyers_pou2f3[order(names(sawyers_pou2f3))]
PRN_pou2f3 <- PRN_pou2f3[order(names(PRN_pou2f3))]

plot_df <- data.frame(sawyers_pou2f3, PRN_pou2f3)

library("ggpubr")
ggscatter(plot_df, x = "sawyers_pou2f3", y = "PRN_pou2f3", 
          add = "reg.line", 
          add.params = list(color = "red", fill = "lightgray"),
          conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "PtRP Pou2f3", ylab = "PRN Pou2f3")


```




## Pseudobullk of Sawyers NEPC cluster
```{r}

sawyers_NEPC <- subset(sawyers, idents = "NEPC")
sawyers_NEPC <- rowMeans(sawyers_NEPC@assays$RNA@data)

names(sawyers_NEPC)  <- tolower(names(sawyers_NEPC))

```

## Pseudobulk of PRN NEPC
```{r}
PRN <- readRDS("./Seurat_Objects/Annotated_PRN_6wk_8wk.rds")
PRN <- SetIdent(PRN, value = "cell.type3")
PRN_NEPC <- subset(PRN, idents = "Neuroendocrine")
rm(PRN)

PRN_NEPC <- rowMeans(PRN_NEPC@assays$RNA@counts)
names(PRN_NEPC)  <- tolower(names(PRN_NEPC))

```

## Spearmans correlation of NEPC clusters
```{r}
overlap <- intersect(names(sawyers_NEPC), names(PRN_NEPC))
sawyers_NEPC <- sawyers_NEPC[names(sawyers_NEPC) %in% overlap]
PRN_NEPC <- PRN_NEPC[names(PRN_NEPC) %in% overlap]

sawyers_NEPC <- sawyers_NEPC[order(names(sawyers_NEPC))]
PRN_NEPC <- PRN_NEPC[order(names(PRN_NEPC))]

plot_df2 <- data.frame(sawyers_NEPC, PRN_NEPC)

library("ggpubr")
ggscatter(plot_df2, x = "sawyers_NEPC", y = "PRN_NEPC", 
          add = "reg.line", 
          add.params = list(color = "red", fill = "lightgray"),
          conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "PtRP NEPC", ylab = "PRN NEPC")


```





## Pseudobulk of PRN T Cells
```{r}
PRN <- readRDS("./Seurat_Objects/Annotated_PRN_6wk_8wk.rds")
PRN <- SetIdent(PRN, value = "cell.type3")
PRN_Granulocytes <- subset(PRN, idents = "Granulocytes")
rm(PRN)

PRN_Granulocytes <- rowMeans(PRN_Granulocytes@assays$RNA@counts)
names(PRN_Granulocytes)  <- tolower(names(PRN_Granulocytes))

```

## Spearmans correlation of Control 
```{r}
overlap <- intersect(names(sawyers_NEPC), names(PRN_Granulocytes))
sawyers_NEPC <- sawyers_NEPC[names(sawyers_NEPC) %in% overlap]
PRN_Granulocytes <- PRN_Granulocytes[names(PRN_Granulocytes) %in% overlap]

sawyers_NEPC <- sawyers_NEPC[order(names(sawyers_NEPC))]
PRN_Granulocytes <- PRN_Granulocytes[order(names(PRN_Granulocytes))]

plot_df3 <- data.frame(sawyers_NEPC, PRN_Granulocytes)

library("ggpubr")
ggscatter(plot_df3, x = "sawyers_NEPC", y = "PRN_Granulocytes", 
          add = "reg.line", 
          add.params = list(color = "red", fill = "lightgray"),
          conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "PtRP NEPC", ylab = "PRN Granulocytes")


```

## Check for IL10 in Pou2f3
```{r}
FeaturePlot(object = sawyers, features = 'IL10', slot = "scale.data")
VlnPlot(sawyers, features = "IL10", slot = "scale.data")
```




## Create Seurat object of full Sawyers anndata with all cell types
```{r}
library(Seurat)
library(SeuratData)
library(SeuratDisk)
library(MuDataSeurat)


sawyers_full <- ReadH5AD("./Sawyers_data/GSM6428979_adata.gemm.combined.final.h5ad")
DefaultAssay(sawyers_full) <- "RNA"
sawyers_full <- SetIdent(sawyers_full, value = "fine_cell_type")
DimPlot(sawyers_full, reduction = "umap")
FeaturePlot(object = sawyers_full, features = 'POU2F3', slot = "scale.data")
FeaturePlot(object = sawyers_full, features = 'IL10', slot = "scale.data")
FeaturePlot(object = sawyers_full, features = 'IL10RA', slot = "scale.data")
FeaturePlot(object = sawyers_full, features = 'IL10RA', slot = "scale.data")

saveRDS(sawyers_full, "./Sawyers_data/Sawyers_seurat_full.rds")
#GSM6428979_adata.gemm.combined.final.h5ad
```


## Subset to only PtRP
```{r}
PtRP <- subset(sawyers_full, genotype == "PRP")
saveRDS(PtRP, "./Sawyers_data/Sawyers_full_PtRP.rds")
```

## Check expression in PtRP
```{r}
PtRP$tp <- factor(x = PtRP$tp, levels = c('8weeks', '9weeks', "12weeks", "16weeks"))
VlnPlot(PtRP, features = c("SOX4", "SOX21", "CTCF"), idents = c("NEPC", "NEPC-P", "EMT", "Mutant_L1", "Mutant_L2", "Mutant_B1", "L3"), slot = "scale.data")
VlnPlot(PtRP, features = c("SOX4", "SOX21", "CTCF"), idents = c("NEPC-P"), slot = "scale.data", group.by = "tp")
VlnPlot(PtRP, features = c("SOX4", "SOX21", "CTCF"), idents = c("NEPC"), slot = "scale.data", group.by = "tp")
VlnPlot(PtRP, features = c("SOX4", "SOX21", "CTCF"), idents = c("NEPC-P"), slot = "scale.data", group.by = "castrate")
VlnPlot(PtRP, features = c("SOX4", "SOX21", "CTCF"), idents = c("NEPC"), slot = "scale.data", group.by = "castrate")
```

